---
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  erro = FALSE,
  message = FALSE,
  comment = "#>"
)
```



# **XCO2**

## **Loading internal functions**

```{r,results='hide'}
functions_files <- list.files('r/',full.names = T)
purrr::map(functions_files,source)
#webshot::install_phantomjs(force=TRUE)
```



## **XCO2 download and pre-processing**

### **Download**
```{r}

## Before ruining this chunk, please create a data-raw folder


# url_filename <- list.files("url/",
#                             pattern = ".txt",
#                             full.names = TRUE)
# 
# urls <- read.table(url_filename) |>
#   dplyr::filter(!stringr::str_detect(V1,".pdf"))
# n_urls <- nrow(urls)
# 
# ### Download
# tictoc::tic()
# furrr::future_pmap(list(urls[,1],"input your user","your password"),my_ncdf4_download)
# tictoc::toc()
```

### **Extraction**

```{r}
# files_names <- list.files("data-raw/",
#                           pattern = "nc",
#                           full.names = TRUE)
# 
# #### Extracting
# 
# xco2 <- purrr::map_df(files_names, my_ncdf4_extractor) |>
#   dplyr::mutate(
#     date = as.Date.POSIXct(time)
#   )
# dplyr::glimpse(xco2)

# ### Saving full data

# readr::write_rds(xco2,'data/xco2_full.rds')
```


### **Aggregation**

```{r}
# Expand grid for brazil
#
##
# dist <- 0.5
# grid_br <- expand.grid(lon=seq(-74,
#                                -27,dist),
#                        lat=seq(-34,
#                                6,
#                                dist))
# plot(grid_br)
# 
# 
# br <- geobr::read_country(showProgress = FALSE)
# region <- geobr::read_region(showProgress = FALSE)
# 
# pol_br <- br$geom |> purrr::pluck(1) |> as.matrix()
# pol_north <- region$geom |> purrr::pluck(1) |> as.matrix()
# pol_northeast <- region$geom |> purrr::pluck(2) |> as.matrix()
# pol_southeast <- region$geom |> purrr::pluck(3) |> as.matrix()
# pol_south <- region$geom |> purrr::pluck(4) |> as.matrix()
# pol_midwest<- region$geom |> purrr::pluck(5) |> as.matrix()
# 
# # correcting poligions
# 
# pol_br <- pol_br[pol_br[,1]<=-34,]
# pol_br <- pol_br[!((pol_br[,1]>=-38.8 & pol_br[,1]<=-38.6) &
#                      (pol_br[,2]>= -19 & pol_br[,2]<= -16)),]
# 
# pol_northeast <- pol_northeast[pol_northeast[,1]<=-34,]
# pol_northeast <- pol_northeast[!((pol_northeast[,1]>=-38.7 &
#                                     pol_northeast[,1]<=-38.6) &
#                                    pol_northeast[,2]<= -15),]
# 
# pol_southeast <- pol_southeast[pol_southeast[,1]<=-30,]
# 
# 
# ### filtering expanded grid to the brazil boundries
# 
# grid_br_cut <- grid_br |>
#   dplyr::mutate(
#     flag_br = def_pol(lon,lat,pol_br),
#     flag_north = def_pol(lon,lat,pol_north),
#     flag_northeast = def_pol(lon,lat,pol_northeast),
#     flag_midwest= def_pol(lon,lat,pol_midwest),
#     flag_southeast = def_pol(lon,lat,pol_southeast),
#     flag_south = def_pol(lon,lat,pol_south)
#   ) |>
#   tidyr::pivot_longer(
#     tidyr::starts_with('flag'),
#     names_to = 'region',
#     values_to = 'flag'
#   ) |>
#   dplyr::filter(flag) |>
#   dplyr::select(lon,lat) |>
#   dplyr::group_by(lon,lat) |>
#   dplyr::summarise(
#     n_obs = dplyr::n()
#   )
# 
# plot(grid_br_cut$lon,grid_br_cut$lat)
# 
# 
# #### aggregation
# xco2df <- readr::read_rds('data/xco2_full.rds')
# 
# xco2_full_trend <- xco2df |> dplyr::mutate(
#   year =lubridate::year(date),
#   month = lubridate::month(date)
# )
# 
# max(xco2_full_trend$year)
# 
# 
# for(i in 2015:2023){
#   aux_xco2 <- xco2_full_trend |>
#     dplyr::filter(year==i)
#   vct_xco2 <- vector();dist_xco2 <- vector();
#   lon_grid <- vector();lat_grid <- vector();
#   for(k in 1:nrow(aux_xco2)){
#     d <- sqrt((aux_xco2$lon[k]-grid_br_cut$lon)^2+
#                 (aux_xco2$lat[k]-grid_br_cut$lat)^2
#     )
#     min_index <- order(d)[1]
#     vct_xco2[k] <- aux_xco2$xco2[min_index]
#     dist_xco2[k] <- d[order(d)[1]]
#     lon_grid[k] <- grid_br_cut$lon[min_index]
#     lat_grid[k] <- grid_br_cut$lat[min_index]
#   }
#   aux_xco2$dist_xco2 <- dist_xco2
#   aux_xco2$xco2_new <- vct_xco2
#   aux_xco2$lon_grid <- lon_grid
#   aux_xco2$lat_grid <- lat_grid
#   if(i == 2015){
#     xco2_full_trend_cut <- aux_xco2
#   }else{
#     xco2_full_trend_cut <- rbind(xco2_full_trend_cut,aux_xco2)
#   }
# }
# 
# 
# xco2_full_trend_cut|>
#   dplyr::mutate(
#     dist_conf = sqrt((lon - lon_grid)^2 + (lat - lat_grid)^2)
#   ) |>
#   dplyr::glimpse()
# 
# nrow(xco2_full_trend_cut |>
#        dplyr::mutate(
#          dist_conf = sqrt((lon - lon_grid)^2 + (lat - lat_grid)^2),
#          dist_bol = dist_xco2 - dist_conf
#        ) |>
#        dplyr::filter(dist_bol ==0)) == nrow(xco2_full_trend_cut)
# 
# ### Saving aggregated data
# readr::write_rds(xco2_full_trend_cut,'data/xco2_0.5deg_full_trend.rds')

```


## **Loading data and shps**

```{r}
br <- geobr::read_country(showProgress = FALSE)
south_file <- list.files('South_America/',pattern = 'shp',full.names = T)
south_america <- sf::read_sf(south_file[1])
biomes <- geobr::read_biomes(showProgress = FALSE) 

xco2df <- readr::read_rds('data/xco2_0.5deg_full_trend.rds')
```

```{r}
biomes2015 <- biomes |> dplyr::mutate(year=2015)
biomes2016 <- biomes |> dplyr::mutate(year=2016)
biomes2017 <- biomes |> dplyr::mutate(year=2017)
biomes2018 <- biomes |> dplyr::mutate(year=2018)
biomes2020 <- biomes |> dplyr::mutate(year=2020)
biomes2021 <- biomes |> dplyr::mutate(year=2021)
biomes2022 <- biomes |> dplyr::mutate(year=2022)

biomes <- rbind(biomes,biomes2015,biomes2016,
                biomes2017,biomes2018,biomes2020,
                biomes2021,biomes2022)

biomes

rm(biomes2015,biomes2016,
                biomes2017,biomes2018,biomes2020,
                biomes2021,biomes2022)
```

## **Data Overview**

### **Descriptive statistics** 

```{r}
df_stat_desc_xco2 <- xco2df |>
  dplyr::filter(dist_xco2<0.25) |>
  dplyr::mutate(
    date = lubridate::make_date(year,month,'15')
  ) |>
  dplyr::group_by(date) |>
  dplyr::summarise(
    N = length(xco2),
    MIN = min(xco2),
    MEAN = mean(xco2),
    MEDIAN = median(xco2),
    MAX = max(xco2),
    VARIANCY  = var(xco2),
    STD_DV = sd(xco2),
    CV = 100*STD_DV/MEAN,
    SKW = agricolae::skewness(xco2),
    KRT = agricolae::kurtosis(xco2),
    )

dplyr::glimpse(df_stat_desc_xco2)
#DT::datatable(df_stat_desc)
```

### **Histograms**

```{r}
xco2df |>
  dplyr::filter(dist_xco2<0.25 & year <2023) |>
  ggplot2::ggplot(ggplot2::aes(x=xco2)) +
  ggplot2::geom_histogram(color="black",fill="gray",
                 bins = 30) +
  ggplot2::facet_wrap(~year, scales = "free") +
  ggplot2::theme_bw()
```

##

```{r}
biomes |> 
  dplyr::filter(name_biome!="Sistema Costeiro") |> 
  ggplot2::ggplot()+
  ggplot2::geom_sf(fill="white")+
  ggplot2::geom_tile(
    data=xco2df |> 
      dplyr::filter(dist_xco2<0.25 & year <2023) |>
      dplyr::group_by(year,lon_grid,lat_grid) |> 
      dplyr::summarise(
        nobs=dplyr::n()
      ) |> #dplyr::pull(nobs) |> max()
      dplyr::mutate(
        nobs_class=dplyr::case_when(
          nobs < 6 ~ 5,
          nobs < 12 ~11,
          nobs < 24 ~ 23,
          nobs < 48 ~ 47,
          nobs<96 ~95,
          nobs<192~191,
          nobs>=192~Inf
        )
      ),
    ggplot2::aes(x=lon_grid,y=lat_grid,
                 fill=forcats::as_factor(nobs_class))
  )+
  ggplot2::scale_fill_viridis_d()+
  #ggplot2::geom_sf(data=biomes,fill=NA)+
  ggplot2::facet_wrap(~year)+
  ggplot2::labs(
    x='longitude',
    y="latitude",
    fill= expression("soundings")
  )
```


```{r}
xco2df |> 
  dplyr::filter(dist_xco2<0.25 & year <2023) |>
  dplyr::group_by(year,lon_grid,lat_grid) |> 
  dplyr::summarise(
    nobs=dplyr::n()
  ) |> 
  #dplyr::arrange(dplyr::desc(nobs)) |> 
  ggplot2::ggplot(ggplot2::aes(x=nobs))+
  ggplot2::geom_histogram()+
  ggplot2::facet_wrap(~year)
  
```


## **Temporal visualization**

### ***General for Brazil***
```{r}

xco2df |>
  dplyr::filter(dist_xco2<0.25) |> # radius (grid cell = 0.5°)
  dplyr::filter(year %in% 2015:2022) |>
  dplyr::group_by(year,date) |>
  dplyr::summarise(xco2_mean=mean(xco2)) |>
  ggplot2::ggplot(ggplot2::aes(x=date,y=xco2_mean))+
  ggplot2::geom_point(shape=21,color="black",fill="gray") +
  ggplot2::geom_line(color="red")+
  ggplot2::geom_smooth(method = "lm") +
  ggplot2::ylim(390,420)+
  ggpmisc::stat_poly_eq(formula = y ~ x, 
               ggplot2::aes(label = paste(..eq.label.., 
                                          ..rr.label.., 
                                 #..p.value.label.., 
                                 sep = "*`,`~")),
               label.y = 0.01,
               parse = TRUE
               )+
  ggplot2::facet_wrap(~year,scales ='free_x')+
  ggplot2::theme_bw()+
  ggplot2::labs(x='',y=expression('Xco'[2]~' (ppm)'),fill='' )

# 
# ggplot2::ggsave('img/xco2_temporal_ano_pvalue.png',units="in", width=10, height=7,
#                 dpi=300)
```

### ***Rationality of beta***
```{r}
xco2df_rationality <- xco2df |>
  dplyr::filter(dist_xco2<0.25) |>
  dplyr::mutate(
    lon = lon_grid,
    lat = lat_grid,
  ) |>
  dplyr::filter(lon == -50, lat ==-19) |>
  dplyr::select(-c(lon_grid,lat_grid)) |>
  dplyr::group_by(lon,lat,year,month) |>
  dplyr::summarise(
    xco2_mean= mean(xco2,na.rm=TRUE),
    xco2_sd = sd(xco2,na.rm=TRUE),
    xco2_uncertanty = mean(uncertanty),
    nobs = dplyr::n(),
    xco2_ep = xco2_sd/sqrt(nobs),
    cv = 100*xco2_sd/xco2_mean
  ) |>
  dplyr::mutate(
    date = lubridate::make_date(year,month,'15')
  )
mod <- lm(xco2_mean~x,data =xco2df_rationality |>
            dplyr::mutate(
              x = 1:dplyr::n()
            ))
xco2df_rationality |>
  dplyr::filter(lubridate::year(date)<2023) |>
  dplyr::mutate(
    x=1:dplyr::n(),
    xco2_est = mod$coefficients[1] + mod$coefficients[2]*x,
    delta=xco2_est - xco2_mean,
    xco2r = (mod$coefficients[1]-delta)-(mean(xco2_mean)-mod$coefficients[1])
  ) |>
  dplyr::group_by(date) |>
  dplyr::summarise(xco2_mean=mean(xco2r)) |>
  ggplot2::ggplot(ggplot2::aes(x=date,y=xco2_mean )) +
  ggplot2::geom_point(shape=21,color="black",fill="gray") +
  ggplot2::geom_line(color="red") +
  ggplot2::geom_smooth(method = "lm") +
  ggpmisc::stat_poly_eq(formula = y ~ x, 
               ggplot2::aes(label = paste(..eq.label.., 
                                          ..rr.label.., 
                                 #..p.value.label.., 
                                 sep = "*`,`~")),
               label.y = 0.01,
               parse = TRUE
               ) +
  ggplot2::facet_wrap(~lubridate::year(date),scales = 'free_x')+
  ggplot2::theme_bw()+
  ggplot2::xlab('Month')+
  ggplot2::ylab(expression(
    'Xco'[2][R]~' (ppm)'
  ))

# ggplot2::ggsave('img/rationality_beta.png',units="in", width=10, height=7,
#                dpi=300)
```




## **Beta analisys and plots on a per-year basis by grid cell**

```{r}
for(i in 2015:2022){
  
  print("===============================================")
  
  print(i)
  
  print("===============================================")
  
  xco2df_filter <- xco2df |>
    dplyr::filter(year == i) |> ## filter by year
    dplyr::filter(dist_xco2<0.25) |>
    dplyr::mutate(
      lon = lon_grid,
      lat = lat_grid,
    ) |>
    dplyr::select(-c(lon_grid,lat_grid)) |>
    dplyr::group_by(lon,lat,year,month) |> ## data month and grid cell aggregation
    dplyr::summarise(
      xco2_mean= mean(xco2,na.rm=TRUE),
      xco2_sd = sd(xco2,na.rm=TRUE),
      xco2_uncertanty = mean(uncertanty),
      nobs = dplyr::n(),
      xco2_ep = xco2_sd/sqrt(nobs),
      cv = 100*xco2_sd/xco2_mean
    ) |>
    dplyr::mutate(
      date = lubridate::make_date(year,month,'15')
    )
  
  ### general model for the year
  
  mod <- lm(xco2_mean~x,data =xco2df_filter |>
              dplyr::mutate(
                x = 1:dplyr::n()
              ))
  
  print("----------------------------")
  print("XCO2 before regionalization")
  
  plot_1 <- xco2df_filter |>
    dplyr::group_by(date) |>
    dplyr::summarise(xco2_mean=mean(xco2_mean)) |>
    ggplot2::ggplot(ggplot2::aes(x=date,y=xco2_mean )) +
    ggplot2::geom_point(shape=21,color="black",fill="gray") +
    ggplot2::geom_line(color="red") +
    ggplot2::geom_smooth(method = "lm") +
    ggpubr::stat_regline_equation(ggplot2::aes(
      label =  paste(..eq.label.., ..rr.label.., sep = "*plain(\",\")~~"))) +
    ggplot2::theme_bw()+
    ggplot2::xlab('Date')+
    ggplot2::ylab(expression(
      'Xco'[2]~' (ppm)'
    ))
  print(plot_1)
  
  #### regionalization of XCO2 
  print("----------------------------")
  print("XCO2 after regionalization")
  
  plot_2 <- xco2df_filter |>
    dplyr::mutate(
      x=1:dplyr::n(),
      xco2_est = mod$coefficients[1] + mod$coefficients[2]*x,
      delta=xco2_est - xco2_mean,
      xco2r = (mod$coefficients[1]-delta)-(mean(xco2_mean)-mod$coefficients[1])
    ) |>
    dplyr::group_by(date) |>
    dplyr::summarise(xco2_mean=mean(xco2r)) |>
    ggplot2::ggplot(ggplot2::aes(x=date,y=xco2_mean )) +
    ggplot2::geom_point(shape=21,color="black",fill="gray") +
    ggplot2::geom_line(color="red") +
    ggplot2::geom_smooth(method = "lm") +
    ggpubr::stat_regline_equation(ggplot2::aes(
      label =  paste(..eq.label.., ..rr.label.., sep = "*plain(\",\")~~"))) +
    ggplot2::theme_bw()+
    ggplot2::xlab('Date')+
    ggplot2::ylab(expression(
      'Xco'[2][R]~' (ppm)'
    ))
  print(plot_2)
  
  
  ### XCO2 regionalized
  xco2detrend <- xco2df_filter |>
    dplyr::mutate(
      x=1:dplyr::n(),
      xco2_est = mod$coefficients[1] + mod$coefficients[2]*x,
      delta=xco2_est - xco2_mean,
      xco2r = (mod$coefficients[1]-delta)-(mean(xco2_mean)-mod$coefficients[1])
    )
  xco2_aux_detrend <- xco2detrend |>
    dplyr::ungroup() |>
    dplyr::group_by(date) |>
    dplyr::summarise(
      xco2 = mean(xco2r)
    )
  
  ## general linear model
  mod_detrend <- lm(xco2 ~date,
                    data = xco2_aux_detrend )
  beta_r <-mod_detrend$coefficients[2] # regional beta
  ep <- summary(mod_detrend)$coefficients[2,2] # regional standard error
  
  ilbr <- beta_r-ep
  slbr <- beta_r+ep
  
  
  ### creating a nest object by grid cell
  xco2_nest <- xco2detrend |>
    tibble::as_tibble() |>
    dplyr::mutate(year =lubridate::year(date),
                  quarter = lubridate::quarter(date),
                  quarter_year = lubridate::make_date(year, quarter, 1)) |>
    dplyr::group_by(lon, lat,date) |>
    dplyr::summarise(xco2 = mean(xco2r, na.rm=TRUE)) |>
    dplyr::mutate(
      id_time = date
    ) |>
    dplyr::group_by(lon,lat) |>
    tidyr::nest()
  
  ### linear regression for each grid cell
  
  xco2_nest_detrend <- xco2_nest|>
    dplyr::mutate(
      beta_line = purrr::map(data,linear_reg, output="beta1"),
      p_value = purrr::map(data,linear_reg, output="p_value"),
      n_obs = purrr::map(data,linear_reg, output="n"),
      beta_error=purrr::map(data,linear_reg,output='betaerror'),
      model_error=purrr::map(data,linear_reg,output='modelerror')
    )
  
  ### creating a table
  
  xco2_aux_detrend_new <- xco2_nest_detrend |>
    dplyr::filter(n_obs > 4) |> ## criteria of minimum observation month for grid cell
    tidyr::unnest(cols = c(beta_line,beta_error,model_error)) |>
    dplyr::ungroup() |>
    dplyr::select(lon, lat, beta_line,beta_error,model_error)
  
  q3_xco2 <- xco2_aux_detrend_new |> dplyr::pull(beta_line) |> quantile(.75)
  q1_xco2 <- xco2_aux_detrend_new |> dplyr::pull(beta_line) |> quantile(.25)
  
  
  plot3 <- xco2_aux_detrend_new |>
    ggplot2::ggplot(ggplot2::aes(x=beta_line)) +
    ggplot2::geom_histogram(bins=30,
                            fill="orange",
                            color="black") +
    ggplot2::labs(x="βpixel",y="Count") +
    ggplot2::geom_vline(xintercept = q1_xco2,
                        color = "red",
                        lty=2) +
    ggplot2::geom_vline(xintercept = q3_xco2,
                        color = "red",
                        lty=2)+
    gghighlight::gghighlight(beta_line < q1_xco2 | beta_line>q3_xco2,
                             unhighlighted_params = list(
                               color = "darkgray",
                               fill = "lightgray")) +
    ggplot2::theme_minimal()
  print("----------------------------")
  print("significant βpixel histogram")
  print(plot3)
  
  plot4 <- south_america |>
    ggplot2::ggplot()+
    ggspatial::annotation_map_tile(type = 'cartolight')+
    ggplot2::geom_sf(col='grey',fill='white')+
    ggplot2::geom_sf(data=br,col='red',fill='NA')+
    ggplot2::ylim(-35,5.5)+
    ggplot2::xlim(-75,-35)+
    ggplot2::geom_tile(data=xco2_aux_detrend_new |>
                         dplyr::mutate(
                           xco2 = dplyr::case_when(
                             beta_line > q3_xco2 ~ 'Source',
                             beta_line < q1_xco2 ~'Sink',
                             .default = 'Non Significant'
                           )
                         ) |>
                         dplyr::filter(xco2!='Non Significant'),
                       ggplot2::aes(x=lon,y=lat,color=xco2,fill=xco2),
    )+
    map_theme_2()+
    ggplot2::scale_color_manual(values = c('darkgreen','darkred'))+
    ggplot2::scale_fill_manual(values = c('darkgreen','darkred'))+
    ggplot2::labs(x='Longitude',y='Latitude',
                  col=expression('Xco'[2]),fill=expression('Xco'[2]))
  print("----------------------------")
  print("significant source and sink")
  print(plot4)
  
  #### CO2 emission and assimilation
  
  plot5 <- south_america |>
    ggplot2::ggplot()+
    ggspatial::annotation_map_tile(type = 'cartolight')+
    ggplot2::geom_sf(col='grey',fill='white')+
    ggplot2::geom_sf(data=br,col='red',fill='NA')+
    ggplot2::ylim(-35,5.5)+
    ggplot2::xlim(-75,-35)+
    ggplot2::geom_tile(data=xco2_aux_detrend_new |>
                         dplyr::mutate(
                           xco2 = dplyr::case_when(
                             beta_line > q3_xco2 ~ 'Source',
                             beta_line < q1_xco2 ~'Sink',
                             .default = 'Non Significant'
                           )
                         ) |>
                         dplyr::filter(xco2!='Non Significant') |>
                         dplyr::mutate(
                           beta_molm=(beta_line*10000)*44/24.45, # beta conversion to mg mol/m3 day * 10000m ===> mg mol / m2 day
                           fco2 = beta_molm*30/1000, # montly fco2 ===> /1000 is to get in grams
                         ),
                       ggplot2::aes(x=lon,y=lat,color=fco2,fill=fco2)
    )+
    ggplot2::scale_color_viridis_c(option = "inferno")+
    ggplot2::scale_fill_viridis_c(option = "inferno")+
    map_theme_2()+
    ggplot2::labs(x='Longitude',y='Latitude',
                  col=expression('FCO'[2]~'(g'~m^-2*month^-1~')'),
                  fill=expression('FCO'[2]~'(g'~m^-2*month^-1~')')
    )
  print("----------------------------")
  print("CO2 Flux")
  print(plot5)
  
  plot6 <- south_america |>
    ggplot2::ggplot()+
    ggspatial::annotation_map_tile(type = 'cartolight')+
    ggplot2::geom_sf(col='grey',fill='white')+
    ggplot2::geom_sf(data=br,col='red',fill='NA')+
    ggplot2::ylim(-35,6)+
    ggplot2::xlim(-75,-35)+
    ggplot2::geom_tile(data=xco2_aux_detrend_new |>
                         dplyr::mutate(
                           xco2 = dplyr::case_when(
                             beta_line > q3_xco2 ~ 'Source',
                             beta_line < q1_xco2 ~'Sink',
                             .default = 'Non Significant'
                           )
                         ) |>
                         dplyr::filter(xco2!='Non Significant')|>
                         dplyr::mutate(
                           beta_molm=(beta_error*10000)*44/24.45,
                           fco2_error = beta_molm*30/1000),
                       ggplot2::aes(x=lon,y=lat,color=fco2_error,fill=fco2_error)
    )+
    ggplot2::scale_color_viridis_c(option = "inferno")+
    ggplot2::scale_fill_viridis_c(option = "inferno")+
    map_theme_2()+
    ggplot2::labs(x='Longitude',y='Latitude',
                  col=expression('FCO'[2]~'error (g'~m^-2*month^-1~')'),
                  fill=expression('FCO'[2]~'error (g'~m^-2*month^-1~')')
    )
  print("----------------------------")
  print("CO2 Flux error")
  print(plot6)
  
  print("===============================================")
  
}

```



## **Creating a unique dataframe**

```{r,results='hide',message=FALSE,warning=FALSE}
for(i in 2015:2022){
  if(i == 2015){
    
    xco2df_filter <- xco2df |>
      dplyr::filter(year == i) |> ## filter by year
      dplyr::filter(dist_xco2<0.25) |>
      dplyr::mutate(
        lon = lon_grid,
        lat = lat_grid,
      ) |>
      dplyr::select(-c(lon_grid,lat_grid)) |>
      dplyr::group_by(lon,lat,year,month) |> ## data month and grid cell aggregation
      dplyr::summarise(
        xco2_mean= mean(xco2,na.rm=TRUE),
        xco2_sd = sd(xco2,na.rm=TRUE),
        xco2_uncertanty = mean(uncertanty),
        nobs = dplyr::n(),
        xco2_ep = xco2_sd/sqrt(nobs),
        cv = 100*xco2_sd/xco2_mean
      ) |>
      dplyr::mutate(
        date = lubridate::make_date(year,month,'15')
      )
    
    ### general model for the year
    
    mod <- lm(xco2_mean~x,data =xco2df_filter |>
                dplyr::mutate(
                  x = 1:dplyr::n()
                ))
    
    
    #### regionalization of XCO2 
    xco2detrend <- xco2df_filter |>
      dplyr::mutate(
        x=1:dplyr::n(),
        xco2_est = mod$coefficients[1] + mod$coefficients[2]*x,
        delta=xco2_est - xco2_mean,
        xco2r = (mod$coefficients[1]-delta)-(mean(xco2_mean)-mod$coefficients[1])
      )
    xco2_aux_detrend <- xco2detrend |>
      dplyr::ungroup() |>
      dplyr::group_by(date) |>
      dplyr::summarise(
        xco2 = mean(xco2r)
      )
    
    ## general linear model
    mod_detrend <- lm(xco2 ~date,
                      data = xco2_aux_detrend )
    beta_r <-mod_detrend$coefficients[2] # regional beta
    ep <- summary(mod_detrend)$coefficients[2,2] # regional standard error
    
    ilbr <- beta_r-ep
    slbr <- beta_r+ep
    
    
    ### creating a nest object by grid cell
    xco2_nest <- xco2detrend |>
      tibble::as_tibble() |>
      dplyr::mutate(year =lubridate::year(date),
                    quarter = lubridate::quarter(date),
                    quarter_year = lubridate::make_date(year, quarter, 1)) |>
      dplyr::group_by(lon, lat,date) |>
      dplyr::summarise(xco2 = mean(xco2r, na.rm=TRUE)) |>
      dplyr::mutate(
        id_time = date
      ) |>
      dplyr::group_by(lon,lat) |>
      tidyr::nest()
    
    ### linear regression for each grid cell
    
    xco2_nest_detrend <- xco2_nest|>
      dplyr::mutate(
        beta_line = purrr::map(data,linear_reg, output="beta1"),
        p_value = purrr::map(data,linear_reg, output="p_value"),
        n_obs = purrr::map(data,linear_reg, output="n"),
        beta_error=purrr::map(data,linear_reg,output='betaerror'),
        model_error=purrr::map(data,linear_reg,output='modelerror')
      )
    
    ### creating a table
    
    xco2_aux_detrend_new <- xco2_nest_detrend |>
      dplyr::filter(n_obs > 4) |> ## criteria of minimum observation month for grid cell
      tidyr::unnest(cols = c(beta_line,beta_error,model_error)) |>
      dplyr::ungroup() |>
      dplyr::select(lon, lat, beta_line,beta_error,model_error) |> 
      dplyr::mutate(year=i)
    
    q3_xco2 <- xco2_aux_detrend_new |> dplyr::pull(beta_line) |> quantile(.75)
    q1_xco2 <- xco2_aux_detrend_new |> dplyr::pull(beta_line) |> quantile(.25)
    
    dfall <- xco2_aux_detrend_new |> dplyr::mutate(
      xco2 = dplyr::case_when(
        beta_line > q3_xco2 ~ 'Source',
        beta_line < q1_xco2 ~'Sink',
        .default = 'Non Significant'
      )
    )
  }else{
    xco2df_filter <- xco2df |>
      dplyr::filter(year == i) |> ## filter by year
      dplyr::filter(dist_xco2<0.25) |>
      dplyr::mutate(
        lon = lon_grid,
        lat = lat_grid,
      ) |>
      dplyr::select(-c(lon_grid,lat_grid)) |>
      dplyr::group_by(lon,lat,year,month) |> ## data month and grid cell aggregation
      dplyr::summarise(
        xco2_mean= mean(xco2,na.rm=TRUE),
        xco2_sd = sd(xco2,na.rm=TRUE),
        xco2_uncertanty = mean(uncertanty),
        nobs = dplyr::n(),
        xco2_ep = xco2_sd/sqrt(nobs),
        cv = 100*xco2_sd/xco2_mean
      ) |>
      dplyr::mutate(
        date = lubridate::make_date(year,month,'15')
      )
    
    ### general model for the year
    
    mod <- lm(xco2_mean~x,data =xco2df_filter |>
                dplyr::mutate(
                  x = 1:dplyr::n()
                ))
    
    
    #### regionalization of XCO2 
    xco2detrend <- xco2df_filter |>
      dplyr::mutate(
        x=1:dplyr::n(),
        xco2_est = mod$coefficients[1] + mod$coefficients[2]*x,
        delta=xco2_est - xco2_mean,
        xco2r = (mod$coefficients[1]-delta)-(mean(xco2_mean)-mod$coefficients[1])
      )
    xco2_aux_detrend <- xco2detrend |>
      dplyr::ungroup() |>
      dplyr::group_by(date) |>
      dplyr::summarise(
        xco2 = mean(xco2r)
      )
    
    ## general linear model
    mod_detrend <- lm(xco2 ~date,
                      data = xco2_aux_detrend )
    beta_r <-mod_detrend$coefficients[2] # regional beta
    ep <- summary(mod_detrend)$coefficients[2,2] # regional standard error
    
    ilbr <- beta_r-ep
    slbr <- beta_r+ep
    
    
    ### creating a nest object by grid cell
    xco2_nest <- xco2detrend |>
      tibble::as_tibble() |>
      dplyr::mutate(year =lubridate::year(date),
                    quarter = lubridate::quarter(date),
                    quarter_year = lubridate::make_date(year, quarter, 1)) |>
      dplyr::group_by(lon, lat,date) |>
      dplyr::summarise(xco2 = mean(xco2r, na.rm=TRUE)) |>
      dplyr::mutate(
        id_time = date
      ) |>
      dplyr::group_by(lon,lat) |>
      tidyr::nest()
    
    ### linear regression for each grid cell
    
    xco2_nest_detrend <- xco2_nest|>
      dplyr::mutate(
        beta_line = purrr::map(data,linear_reg, output="beta1"),
        p_value = purrr::map(data,linear_reg, output="p_value"),
        n_obs = purrr::map(data,linear_reg, output="n"),
        beta_error=purrr::map(data,linear_reg,output='betaerror'),
        model_error=purrr::map(data,linear_reg,output='modelerror')
      )
    
    ### creating a table
    
    xco2_aux_detrend_new <- xco2_nest_detrend |>
      dplyr::filter(n_obs > 4) |> ## criteria of minimum observation month for grid cell
      tidyr::unnest(cols = c(beta_line,beta_error,model_error)) |>
      dplyr::ungroup() |>
      dplyr::select(lon, lat, beta_line,beta_error,model_error) |> 
      dplyr::mutate(year=i)
    
    q3_xco2 <- xco2_aux_detrend_new |> dplyr::pull(beta_line) |> quantile(.75)
    q1_xco2 <- xco2_aux_detrend_new |> dplyr::pull(beta_line) |> quantile(.25)
    
    dfall_aux <- xco2_aux_detrend_new |> dplyr::mutate(
      xco2 = dplyr::case_when(
        beta_line > q3_xco2 ~ 'Source',
        beta_line < q1_xco2 ~'Sink',
        .default = 'Non Significant'
      )
    )
    
    dfall <- rbind(dfall,dfall_aux)
  }
}


#### saving the file

# writexl::write_xlsx(dfall |>
#                       dplyr::filter(xco2!='NonSignificant') |>
#                       dplyr::mutate(
#                          beta_molm=(10000*beta_line)*44/24.45,
#                          beta_fco2 = beta_molm*30/1000,
#                          beta_molm_erro=(10000*beta_error)*44/24.45,
#                          betaerror_fco2 = beta_molm_erro*30/1000
#                        ),
#                     "output/beta_significant.xlsx")

```


### **Plotting all beta**
```{r}
dfall |> 
  ggplot2::ggplot(ggplot2::aes(x=beta_line*30, y=as.character(year),
                               col=as.character(year), fill=as.character(year)))+
  ggridges::geom_density_ridges(alpha=.2)+
  ggplot2::geom_vline(xintercept =dfall |>
                        dplyr::summarise(b=median(beta_line)) |>
                        dplyr::pull(b)*30, linetype='dashed', color='black')+
  ggplot2::scale_color_viridis_d()+
  ggplot2::scale_fill_viridis_d()+
  #ggplot2::facet_wrap(~xco2,scales = 'free_x')+
  ggplot2::theme_bw()+
  ggplot2::labs(x=expression('βxco'[2]~'(ppm month'^-1~')'),y='',col='',fill='')
```


### **Avegerage CO2 balance for Brazil**
```{r}
dfall |>
  dplyr::filter(xco2!='Non Significant') |> 
  dplyr::mutate(
    beta_molm=(10000*beta_line)*44/24.45, ## ppm to mg/m2 day (10000m is the column)
    beta_fco2 = beta_molm*30/1000, ## mg mol/m2 day ===> g CO2 month^-1 
    beta_molm_erro=(10000*beta_error)*44/24.45,
    betaerror_fco2 = beta_molm_erro*30/1000
  ) |> 
  dplyr::group_by(year) |>
  dplyr::summarise(
    fco2=mean(beta_fco2),
    fco2err=mean(betaerror_fco2),
    effective_balance = sum(12*beta_fco2*(55000^2))/1e15, ## 12 is month 55000^2 is the grid cell area in m^2 
    mean_balance = 12*fco2*8510000000000/1e15,## 12 is month 8.51e12 is brazil's area in m^2 
    mean_error = 12*fco2err*8510000000000/1e15
  ) |>
  ggplot2::ggplot(ggplot2::aes(x=year,y=mean_balance,
                               ymax=mean_balance+mean_error,
                               ymin=mean_balance))+
  ggplot2::geom_col()+
  ggplot2::geom_errorbar()+
  ggplot2::labs(
    x='',
    y=expression('CO'[2]~' Balance (Pg)')
  )+
  ggplot2::theme_bw()
```


### **All CO2 sources and sinks**
```{r}
biomes |>
  dplyr::filter(name_biome!="Sistema Costeiro") |> 
  ggplot2::ggplot()+
  ggspatial::annotation_map_tile(type = "cartolight")+
  ggplot2::geom_sf(data=south_america,col='grey',fill='white')+
  ggplot2::geom_sf(col='black',fill='grey50')+
  ggplot2::ylim(-35,5.5)+
  ggplot2::xlim(-75,-35)+
  ggplot2::geom_tile(data=dfall |>
                       dplyr::filter(xco2!='Non Significant'),
                     ggplot2::aes(x=lon,y=lat,color=xco2,fill=xco2),
  )+
  ggplot2::facet_wrap(~year)+
  ggplot2::theme(axis.text = ggplot2::element_text(size=12),
                 axis.title = ggplot2::element_text(size=20),
                 text = ggplot2::element_text(size=20)
                 )+
  map_theme_2()+
  ggplot2::theme(legend.position = c(1,0),
                 legend.justification = c(1, 0))+
  ggplot2::scale_color_manual(values = c('green','red'))+
  ggplot2::scale_fill_manual(values = c('green','red'))+
  ggplot2::labs(x='Longitude',y='Latitude',
                col=expression('Xco'[2]),fill=expression('Xco'[2]))

# ggplot2::ggsave('img/xco2_identification_all_teste.png',units="in",
#                 width=11, height=11,
#                 dpi=300)
```

### **All FCO2**

```{r}
biomes |>
  dplyr::filter(name_biome!="Sistema Costeiro") |> 
  ggplot2::ggplot()+
  ggspatial::annotation_map_tile(type = "cartolight")+
  ggplot2::geom_sf(data=south_america,col='grey',fill='white')+
  ggplot2::geom_sf(col='black',fill='grey50')+
  ggplot2::ylim(-35,5.5)+
  ggplot2::xlim(-75,-35)+
  ggplot2::geom_tile(data=dfall |>
                       dplyr::filter(xco2!='Non Significant')|> 
                       dplyr::mutate(
                         beta_molm=(10000*beta_line)*44/24.45, 
                         beta_fco2 = beta_molm*30/1000, 
                         beta_molm_erro=(10000*beta_error)*44/24.45,
                         betaerror_fco2 = beta_molm_erro*30/1000
                       ), 
                     ggplot2::aes(x=lon,y=lat,
                                  color=beta_fco2,fill=beta_fco2),
  )+
  ggplot2::facet_wrap(~year)+
  ggplot2::theme(axis.text = ggplot2::element_text(size=12),
                 axis.title = ggplot2::element_text(size=20),
                 text = ggplot2::element_text(size=20)
  )+
  map_theme_2()+
  ggplot2::theme(legend.position = c(1,0),
                 legend.justification = c(1, 0))+
  ggplot2::scale_color_viridis_c(option = "inferno")+
  ggplot2::scale_fill_viridis_c(option = "inferno")+
  ggplot2::labs(x='Longitude',y='Latitude',
                col=expression('FCO'[2]),
                fill=expression('FCO'[2])
  )

# ggplot2::ggsave('img/xco2_fco2_all.png',units="in", width=11, height=11,
#                 dpi=300)
```

### **All FCO2 errors**
```{r}
biomes |>
  dplyr::filter(name_biome!="Sistema Costeiro") |> 
  ggplot2::ggplot()+
  ggspatial::annotation_map_tile(type = "cartolight")+
  ggplot2::geom_sf(data=south_america,col='grey',fill='white')+
  ggplot2::geom_sf(col='black',fill='grey50')+
  ggplot2::ylim(-35,5.5)+
  ggplot2::xlim(-75,-35)+
  ggplot2::geom_tile(data=dfall |>
                       dplyr::filter(xco2!='Non Significant')|> 
                       dplyr::mutate(
                         beta_molm=(10000*beta_line)*44/24.45, 
                         beta_fco2 = beta_molm*30/1000, 
                         beta_molm_erro=(10000*beta_error)*44/24.45,
                         betaerror_fco2 = beta_molm_erro*30/1000
                       ), 
                     ggplot2::aes(x=lon,y=lat,color=betaerror_fco2,fill=betaerror_fco2),
  )+
  ggplot2::facet_wrap(~year)+
  ggplot2::theme(axis.text = ggplot2::element_text(size=12),
                 axis.title = ggplot2::element_text(size=20),
                 text = ggplot2::element_text(size=20)
  )+
  map_theme_2()+
  ggplot2::theme(legend.position = c(1,0),
                 legend.justification = c(1, 0))+
  ggplot2::scale_color_viridis_c(direction = -1)+
  ggplot2::scale_fill_viridis_c(direction=-1)+
  ggplot2::labs(x='Longitude',y='Latitude',
                col=expression(epsilon~'FCO'[2]),
                fill=expression(epsilon~'FCO'[2])
  )
ggplot2::ggsave('img/xco2_fco2_erro_all.png',units="in",
                width=11,height=11,
                dpi=300)
```

# **SIF**

## **Loading internal functions**

```{r,results='hide'}
functions_files <- list.files('r/',full.names = T)
purrr::map(functions_files,source)
#webshot::install_phantomjs(force=TRUE)
```



## **SIF download and pre-processing**

### **Download**
```{r}
## Before ruining this chunk, please create a data-raw folder on SIF folder
### SIF/data-raw/

# url_filename <- list.files("SIF/url/",
#                             pattern = ".txt",
#                             full.names = TRUE)
# 
# urls <- read.table(url_filename) |>
#   dplyr::filter(!stringr::str_detect(V1,".pdf"))
# n_urls <- nrow(urls)
# 
# ### Download
# tictoc::tic()
# furrr::future_pmap(list(urls[,1],"input your user","your password"),my_sif_download)
# tictoc::toc()
```

### **Extraction**

```{r}
# files_names <- list.files("SIF/data-raw/",
#                           pattern = "nc",
#                           full.names = TRUE)
# 
# #### Extracting
# 
# sif_df <- purrr::map_df(files_names, my_sif_extractor)
# dplyr::glimpse(xco2)
# 
# readr::write_rds(sif_df,'SIF/data/sif_full.rds')
```


### **Aggregation**

```{r}
# Expand grid for brazil
#
##
# dist <- 0.5
# grid_br <- expand.grid(lon=seq(-74,
#                                -27,dist),
#                        lat=seq(-34,
#                                6,
#                                dist))
# plot(grid_br)
# 
# 
# br <- geobr::read_country(showProgress = FALSE)
# region <- geobr::read_region(showProgress = FALSE)
# 
# pol_br <- br$geom |> purrr::pluck(1) |> as.matrix()
# pol_north <- region$geom |> purrr::pluck(1) |> as.matrix()
# pol_northeast <- region$geom |> purrr::pluck(2) |> as.matrix()
# pol_southeast <- region$geom |> purrr::pluck(3) |> as.matrix()
# pol_south <- region$geom |> purrr::pluck(4) |> as.matrix()
# pol_midwest<- region$geom |> purrr::pluck(5) |> as.matrix()
# 
# # correcting poligions
# 
# pol_br <- pol_br[pol_br[,1]<=-34,]
# pol_br <- pol_br[!((pol_br[,1]>=-38.8 & pol_br[,1]<=-38.6) &
#                      (pol_br[,2]>= -19 & pol_br[,2]<= -16)),]
# 
# pol_northeast <- pol_northeast[pol_northeast[,1]<=-34,]
# pol_northeast <- pol_northeast[!((pol_northeast[,1]>=-38.7 &
#                                     pol_northeast[,1]<=-38.6) &
#                                    pol_northeast[,2]<= -15),]
# 
# pol_southeast <- pol_southeast[pol_southeast[,1]<=-30,]
# 
# 
# ### filtering expanded grid to the brazil boundries
# 
# grid_br_cut <- grid_br |>
#   dplyr::mutate(
#     flag_br = def_pol(lon,lat,pol_br),
#     flag_north = def_pol(lon,lat,pol_north),
#     flag_northeast = def_pol(lon,lat,pol_northeast),
#     flag_midwest= def_pol(lon,lat,pol_midwest),
#     flag_southeast = def_pol(lon,lat,pol_southeast),
#     flag_south = def_pol(lon,lat,pol_south)
#   ) |>
#   tidyr::pivot_longer(
#     tidyr::starts_with('flag'),
#     names_to = 'region',
#     values_to = 'flag'
#   ) |>
#   dplyr::filter(flag) |>
#   dplyr::select(lon,lat) |>
#   dplyr::group_by(lon,lat) |>
#   dplyr::summarise(
#     n_obs = dplyr::n()
#   )
# 
# plot(grid_br_cut$lon,grid_br_cut$lat)
# 
# 
# #### aggregation


# sifdf <- readr::read_rds('SIF/data/sif_full.rds')
# 
# sif_full <- sifdf |> dplyr::mutate(
#   date=lubridate::as_datetime(time,origin='1990-01-01 00:00:00 UTC'),
#   date = lubridate::as_date(date),
#   year =lubridate::year(date),
#   month = lubridate::month(date)
# )
# 
# max(sif_full$year)
# 
# 
# for(i in 2015:2023){
#   aux_sif <- sif_full |>
#     dplyr::filter(year==i)
#   vct_sif <- vector();dist_sif <- vector();
#   lon_grid <- vector();lat_grid <- vector();
#   for(k in 1:nrow(aux_sif)){
#     d <- sqrt((aux_sif$lon[k]-grid_br_cut$lon)^2+
#                 (aux_sif$lat[k]-grid_br_cut$lat)^2
#     )
#     min_index <- order(d)[1]
#     vct_sif[k] <- aux_sif$sif_757[min_index]
#     dist_sif[k] <- d[order(d)[1]]
#     lon_grid[k] <- grid_br_cut$lon[min_index]
#     lat_grid[k] <- grid_br_cut$lat[min_index]
#   }
#   aux_sif$dist_sif<- dist_sif
#   aux_sif$sif_new <- vct_sif
#   aux_sif$lon_grid <- lon_grid
#   aux_sif$lat_grid <- lat_grid
#   if(i == 2015){
#     sif_full_cut <- aux_sif
#   }else{
#     sif_full_cut <- rbind(sif_full_cut,aux_sif)
#   }
# }
# 
# 
# sif_full_cut|>
#   dplyr::mutate(
#     dist_conf = sqrt((lon - lon_grid)^2 + (lat - lat_grid)^2)
#   ) |>
#   dplyr::glimpse()
# 
# nrow(sif_full_cut |>
#        dplyr::mutate(
#          dist_conf = sqrt((lon - lon_grid)^2 + (lat - lat_grid)^2),
#          dist_bol = dist_sif - dist_conf
#        ) |>
#        dplyr::filter(dist_bol ==0)) == nrow(sif_full_cut)
# 
# 
# readr::write_rds(sif_full_cut,'SIF/data/sif_0.5deg_full_trend.rds')

```


## **Loading data and shps**

```{r}
br <- geobr::read_country(showProgress = FALSE)
south_file <- list.files('South_America/',pattern = 'shp',full.names = T)
south_america <- sf::read_sf(south_file[1])
biomes <- geobr::read_biomes(showProgress = FALSE) 

sifdf <- readr::read_rds('SIF/data/sif_0.5deg_full_trend.rds')
```


## **Data Overview**

### **Descriptive statistics** 

```{r}
df_stat_desc_sif <- sifdf |>
  dplyr::filter(dist_sif<0.25) |>
  dplyr::mutate(
    date = lubridate::make_date(year,month,'15')
  ) |>
  dplyr::group_by(date) |>
  dplyr::summarise(
    N = length(sif_757),
    MIN = min(sif_757),
    MEAN = mean(sif_757),
    MEDIAN = median(sif_757),
    MAX = max(sif_757),
    VARIANCY  = var(sif_757),
    STD_DV = sd(sif_757),
    CV = 100*STD_DV/MEAN,
    SKW = agricolae::skewness(sif_757),
    KRT = agricolae::kurtosis(sif_757),
    )

dplyr::glimpse(df_stat_desc_sif)
#DT::datatable(df_stat_desc)
```

### **Histograms**

```{r}
sifdf |>
  dplyr::filter(dist_sif<0.25 & year <2023) |>
  ggplot2::ggplot(ggplot2::aes(x=sif_757)) +
  ggplot2::geom_histogram(color="black",fill="gray",
                 bins = 30) +
  ggplot2::facet_wrap(~year, scales = "free") +
  ggplot2::theme_bw()
```



## **Temporal visualization**

### ***General for Brazil***
```{r}

sifdf |>
  dplyr::filter(dist_sif<0.25) |> # radius (grid cell = 0.5°)
  dplyr::filter(year %in% 2015:2022) |>
  dplyr::group_by(year,date) |>
  dplyr::summarise(sif=mean(sif_757)) |>
  ggplot2::ggplot(ggplot2::aes(x=date,y=sif))+
  ggplot2::geom_point(shape=21,color="black",fill="gray") +
  ggplot2::geom_line(color="red")+
  ggplot2::geom_smooth(method = "lm") +
  #ggplot2::ylim(390,420)+
  ggpubr::stat_regline_equation(ggplot2::aes(
                                  label =  paste(..eq.label.., ..rr.label..,
                                                 sep = "*plain(\",\")~~")),
                               # label.y = 420
                               ) +
  #ggplot2::facet_wrap(~year,scales ='free')+
  ggplot2::theme_bw()+
  ggplot2::labs(x='',y=expression('SIF 757nm (Wm'^-2~'sr'^-1~mu~'m'^-1~')'),fill='' )
```


## **Pré-processing**
```{r}

sifdf_filter <- sifdf |>
  #dplyr::filter(year == 2015) |>
  dplyr::filter(dist_sif<0.25) |>
  dplyr::mutate(
    lon = lon_grid,
    lat = lat_grid,
  ) |>
  dplyr::select(-c(lon_grid,lat_grid)) |>
  dplyr::group_by(lon,lat,year,month) |>
  dplyr::summarise(
    sif_mean= mean(sif_757,na.rm=TRUE),
    sif_sd = sd(sif_757,na.rm=TRUE),
    sza = mean(sza),
    nobs = dplyr::n(),
    sif_ep = sif_sd/sqrt(nobs),
    cv = 100*sif_sd/sif_mean
  ) |>
  dplyr::mutate(
    date = lubridate::make_date(year,month,'15')
  )

### SIF medians for the period in per grid cell basis
sif_medians <- sifdf_filter |>
  dplyr::group_by(lon,lat) |>
  dplyr::summarise(
    sif_median = median(sif_mean)
  )
```



## **SIF normalized**

```{r}
biomes |>
  dplyr::filter(name_biome!="Sistema Costeiro") |> 
  ggplot2::ggplot()+
  ggspatial::annotation_map_tile(type = "cartolight")+
  ggplot2::geom_sf(data=south_america,col='grey',fill='white')+
  ggplot2::geom_sf(col='black',fill='grey50')+
  ggplot2::ylim(-35,5.5)+
  ggplot2::xlim(-75,-35)+
  ggplot2::geom_tile(data=
                       sifdf_filter |>
                       dplyr::select(lon, lat, date,sif_mean) |>
                       dplyr::left_join(sif_medians) |>
                       dplyr::group_by(lon,lat) |>
                       dplyr::mutate(
                         sif_norm= (sif_mean - sif_median)/sif_median,
                         year=lubridate::year(date)
                       ) |>
                       dplyr::group_by(lon,lat,year) |>
                       dplyr::summarise(
                         sif=mean(sif_norm)
                       ) |>
                       dplyr::mutate(
                         sif_class = ifelse(sif<=0,'Decreasing', 'Enhancement')
                       ) |> dplyr::filter(year!=2023),
                     ggplot2::aes(x=lon,y=lat,fill=sif_class),
  )+
  ggplot2::geom_sf(col='black',fill=NA)+
  ggplot2::facet_wrap(~year)+
  ggplot2::theme(axis.text = ggplot2::element_text(size=12),
                 axis.title = ggplot2::element_text(size=20),
                 text = ggplot2::element_text(size=20))+
  map_theme_2()+
  #ggplot2::scale_fill_viridis_d(option='inferno')+
  ggplot2::scale_fill_manual(values = c('red','green'))+
  ggplot2::theme(legend.position = c(1,0),
                 legend.justification = c(1, 0))+
  ggplot2::labs(x='Longitude',y='Latitude',fill=expression('SIF'[Norm]))

# ggplot2::ggsave('img/sif_norm_all_viridis.png',units="in", 
#                 width=11, height=11,
#                 dpi=300)
```


# **Relationship between SIF norm and Beta**

```{r}
sif_proce <- sifdf_filter |>
  dplyr::select(lon, lat, date,sif_mean) |>
  dplyr::left_join(sif_medians) |>
  dplyr::group_by(lon,lat) |>
  dplyr::mutate(
    sif_norm= (sif_mean - sif_median)/sif_median,
    year=lubridate::year(date)
  ) |>
  dplyr::group_by(lon,lat,year) |>
  dplyr::summarise(
    sif=mean(sif_norm)
  ) |>
  dplyr::mutate(
    sif_class = ifelse(sif<=0,'Decreasing', 'Enhancement')
  )


## read beta data frame if is not in the environment 

# dfall <- readxl::read_excel("output/beta_significant.xlsx")

```


```{r}
my_phi_matrix <-dfall |> 
  dplyr::filter(xco2!="Non Significant") |> 
  dplyr::select(lon,lat,year,beta_line,xco2) |> 
  dplyr::left_join(sif_proce) |> 
  na.omit() |> 
  dplyr::group_by(lon, lat) |> 
  dplyr::mutate(
    new_class = dplyr::case_when(
      xco2 == "Source" & sif_class == "Enhancement" ~ "PP",
      xco2 == "Source" & sif_class == "Decreasing" ~ "PN",
      xco2 == "Sink" & sif_class == "Enhancement" ~ "NP",
      xco2 == "Sink" & sif_class == "Decreasing" ~ "NN"
    )
  ) |> 
  dplyr::group_by(lon,lat, new_class) |> 
  dplyr::summarise(
    nobs = dplyr::n()
  )|> 
  tidyr::pivot_wider(
    names_from = new_class,
    values_from = nobs,
    values_fill = 0
  ) |> 
  dplyr::mutate(
    total = sum(NN,PP,NP,PN)
  ) |> 
  dplyr::filter(total >2) |> 
  dplyr::mutate(
    ac = PP+NP,
    bd = PN+NN,
    ab = PP+PN,
    cd = NP+NN,
    ad = PP*NN,
    bc = PN*NP,
    nume = ad-bc,
    deno = sqrt((ab*ac*bd*cd)),
    phi = ifelse(deno !=0, nume/deno,0)
  )  |> 
  na.omit()

my_phi_matrix |> dplyr::glimpse()
```



```{r}
biomes |>
  dplyr::filter(name_biome!="Sistema Costeiro") |> 
  ggplot2::ggplot()+
  ggspatial::annotation_map_tile(type = "cartolight")+
  ggplot2::geom_sf(data=south_america,col='grey',fill='white')+
  ggplot2::geom_sf(col='black',fill='grey50')+
  ggplot2::ylim(-35,5.5)+
  ggplot2::xlim(-75,-35)+ 
  # ggplot2::ggplot()+
  # ggplot2::geom_sf(fill="white", color="grey10",
  #                  size=.15, show.legend = FALSE)+
  ggplot2::geom_tile(data = my_phi_matrix,
                     ggplot2::aes(y=lat,x=lon,fill=phi)
  )+
  map_theme_2()+
  ggplot2::scale_fill_gradientn(colors=c("darkmagenta","navyblue",
                                      "white","darkorange","darkred"),
                                breaks=c(-1,-0.5,0,0.5,1))+
  # ggplot2::scale_fill_viridis_c(limits=c(-1,1),
  #                               option='inferno')+
  ggplot2::labs(x='Longitude',y='Latitude',fill=expression(phi))

# ggplot2::ggsave('img/phi_map.png',units="in", width=11, height=11,
#                 dpi=300)
```
####


```{r}
dfall |> 
  dplyr::filter(xco2!="Non Significant") |> 
  dplyr::select(lon,lat,year,beta_line,xco2) |> 
  dplyr::left_join(sif_proce) |> 
  na.omit() |> 
  dplyr::group_by(lon, lat) |> 
  dplyr::mutate(
    new_class = dplyr::case_when(
      xco2 == "Source" & sif_class == "Enhancement" ~ "PP",
      xco2 == "Source" & sif_class == "Decreasing" ~ "PN",
      xco2 == "Sink" & sif_class == "Enhancement" ~ "NP",
      xco2 == "Sink" & sif_class == "Decreasing" ~ "NN"
    )
  ) |> 
  dplyr::group_by(new_class) |> 
  dplyr::summarise(
    nobs = dplyr::n()
  ) |> 
  tidyr::pivot_wider(
    names_from = new_class,
    values_from = nobs,
    values_fill = 0
  ) |> 
  dplyr::mutate(
    ratio = NP/(NP+NN)
  )
```

### **linear regression by month for brazil**

```{r}
mod <- lm(xco2_mean~x,data =df_stat_desc_xco2 |>
            dplyr::mutate(
              xco2_mean = MEAN,
              x = 1:dplyr::n()
            ))
df_stat_desc_xco2 |>
  dplyr::filter(lubridate::year(date)<2023) |>
  dplyr::mutate(
    xco2_mean = MEAN,
    x=1:dplyr::n(),
    xco2_est = mod$coefficients[1] + mod$coefficients[2]*x,
    delta=xco2_est - xco2_mean,
    xco2r = (mod$coefficients[1]-delta)-(mean(xco2_mean)-mod$coefficients[1])
  ) |>
  dplyr::group_by(date) |>
  dplyr::summarise(xco2_mean=mean(xco2r)) |>
  dplyr::left_join(df_stat_desc_sif |> 
                     dplyr::select(date,MEAN) |> 
                     dplyr::rename(sif=MEAN)) |> 
  ggplot2::ggplot(ggplot2::aes(x=sif,y=xco2_mean))+
  ggplot2::geom_point()+
  ggplot2::geom_smooth(method = "lm")+
  ggpubr::stat_regline_equation(ggplot2::aes(
    label =  paste(..eq.label.., ..rr.label.., sep = "*plain(\",\")~~"))) +
  ggplot2::theme_bw()+
  ggplot2::theme(
    axis.text = ggplot2::element_text(color='black'),
    axis.ticks = ggplot2::element_line(color='black')
  )+
  ggplot2::xlab(expression(
    'SIF 757nm (Wm'^-2~'sr'^-1~mu~'m'^-1~')'))+
  ggplot2::ylab(expression(
    'Xco'[2][R]~' (ppm)'
  ))
  

```





# **Comparison with other studies**

```{r}
df_comparison <- readxl::read_excel('comparison/comparison.xlsx')

df_comparison |> 
  dplyr::group_by(Source,Biome) |> 
  dplyr::summarise(
    FCO2=mean(FCO2,na.rm=T),
    std=mean(std,na.rm=T)
  ) |> 
  ggplot2::ggplot(ggplot2::aes(x=Source,y=FCO2,
                               ymax=FCO2+std,
                               ymin=FCO2-std,
                               col=Source
  ))+
  ggplot2::geom_point(fill='white', alpha=.5)+
  ggplot2::geom_errorbar()+
  ggplot2::facet_wrap(~Biome,scales = 'free')+
  ggplot2::scale_color_viridis_d()+
  ggplot2::labs(x='',
                y=expression('FCO'[2]~'(g CO'[2]~m^-2*month^-1~')'))+
  ggplot2::theme(axis.text.x  = ggplot2::element_text(hjust = 1,angle = 45))


# ggsave('figure.png',
#        units="in", width=12, height=6,
#        dpi=300)

```


```{r}
df_comparison |> 
  dplyr::group_by(Source,Biome) |> 
  dplyr::summarise(
    FCO2=mean(FCO2,na.rm=T),
    std=mean(std,na.rm=T)
  )
```

